---
title: "p8105_hw1_asf2164"
output: html_document
date: "2022-09-20"
---

Homework 1

Abigail Feinberg

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
# install.packages("palmerpenguins")
```

## R Markdown
```{r, results = "hide"}
data("penguins", package = "palmerpenguins")
# variables: species, island, bill_length_mm, bill_depth_mm, flipper_lenght_mm, body_mass_mm, body_mass_g, sex, year
summary(penguins)
nrow(penguins) # rows = 344
ncol(penguins) # columns = 8


library(tidyr)
drop_na(penguins, flipper_length_mm)
mean(penguins$flipper_length_mm, na.rm = TRUE) # mean of flipper_length_mm = 200.9152


ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguin_plot.pdf")


```

There are `r nrow(penguins)` entries and `r ncol(penguins)` variables in this dataset. There variables that were measured are: 

* species

* island

* bill_length_mm

* bill_depth_mm 

* flipper_lenght_mm

* body_mass_mm 

* body_mass_g

* sex

* year

The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm= TRUE)`. 

## Problem 2 
```{r, results = 'hide'}
set.seed(3)

problem2plot = tibble(
  normal_var = rnorm(10, mean = 0, sd = 1),
  logical_var = normal_var > 0,
  char_var = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  factor_var = factor(
    c(
      "small",
      "medium",
      "large",
      "large",
      "large",
      "medium",
      "medium",
      "small",
      "large",
      "small"
    )
  )
)

mean(pull(problem2plot, var = normal_var)) #mean = -0.06713568
mean(pull(problem2plot, var = logical_var)) #mean = 0.6
mean(pull(problem2plot, var = char_var)) #mean = N/A, this argument is not numerical or logical so means were not calculated
mean(pull(problem2plot, var = factor_var)) #mean = N/A, this argument is not numerical or logical so means were not calculated 

as.numeric(problem2plot$normal_var) #[1] -0.96193342 -0.29252572  0.25878822 -1.15213189  0.19578283  0.03012394  0.08541773  1.11661021 -1.21885742 [10]  1.26736872
as.numeric(problem2plot$logical_var) # [1] 0 0 1 0 1 1 1 1 0 1
as.numeric(problem2plot$char_var) #[1] NA NA NA NA NA NA NA NA NA NA, gives warning that N/As were introduced by coercion, this variable cannot produce a numeric value as a character variable 
as.numeric(problem2plot$factor_var) #[1] 3 2 1 1 1 2 2 3 1 3, factor variables are ordinal. even though these values are understood as numbers by R, you cannot take a mean of these values because they're discrete values 

mean(as.numeric(pull(problem2plot, factor_var))) #mean = 1.9, factor variables are ordinal. even though these values are understood as numbers by R, you cannot take a mean of these values because they're discrete values. this value is meaningless. 

```

The means that work are for the random standard Normal distribution variable and the logical variable. The means that don't work are the character variable and factor variable. This is because the first 2 variables are numeric variables and the second 2 are character variables. 

For as.numeric, the function produces numeric values for all of the variables except the character variable. This is because all of the variables have an order to them and can be converted into numeric values except for the character variables with any set order. However, just because you can take means for all of the variables except the character variables, you should not because the mean value is only meaningful for the random standard normal distribution variable. All of the other variable produce discrete values, not continuous. 
